---
title: "2021_11_09_lecutre03"
author: "Svenja Dobelmann"
date: "9 11 2021"
output: html_document
---


## funny quotes:
```{r}
library(fortunes)
fortune()
fortune("memory")
fortune("data")
fortune("statistics")
```
```{r}
library(cowsay)
say("Hello World!")#

say_hello<- function(){
  animal<- sample(names(animals),1)
  say(paste("Hello, I´m a " , animal,"",collapse = ""), by=animal,what_color="green",length = 25)
}
say_hello()
```
```{r}
names(animals)
```
```{r}
my_fortune<- function(){
  animal<- sample(names(animals),1)
  say(paste(fortune(),collapse="\n"),by=animal)
}
my_fortune()
```
##brickr package for lego style
```{r}
install.packages("brickr")
library(brickr)

bricks_from_table(brick)
brick <- data.frame(
 Level="A",
 X1 = rep(3,4), #The number 3 is the brickrID for 'bright red'
 X2 = rep(3,4)
)
```


always wanted to know what your whatsapp conversation is about and who send what? -> rwhatsapp


## important note:  
save scripts after course
this allows to quickly execute prevoiusly created scripts
think about 10 finger typing! 


## ggplot2 Introduction: 

why? 
- robust, versatile for aesthetic plots
consistent underlying grammar of graphics
plot specification 
very flexible
theme system for polishing plot apprearance
mature and complete graphics

plot = data + aesthetic + geometrie 
-data to visualize
-how to display the data
-type of graphics

other: faceting
annotation
statistics
scales 

### grammar: 
data 
as
ge
stat
scales
facets
coord system 
position adjustment 

```{r}
install.packages("ggplot2")
library(ggplot2)

x <- data.frame(x=1,y=1,label="ggplot2 Introduction \n@ EAGLE")
x11()
 ggplot(data= x, aes(x=x,y=y))+ geom_text(aes(label=label),size=10)
```
### create some data for plotting examples: 
```{r}
x1 <- rnorm(1000,0,1)
x2 <- rnorm(1000,5,10)
x3 <- rep(c("catA","catB","catC","catC","catC","catC"),200)[1:1000]
x4 <- factor(rep(c("yes","no"),500))
df <- data.frame(a=x1,b=x2,c=x3,d=x4)
```


```{r}
ggplot(df,aes(a,b))
```
just an empty plot, in needs a geom
```{r}
ggplot(df,aes(a,b))+geom_point()
```
adding color from "c" column: 

```{r}
ggplot(df,aes(a,b,color=c))+geom_point(alpha=0.5)+ labs(title="first plot",x="x axis \n and a new line")
```
```{r}
ggplot(df,aes(a)) + geom_histogram(fill="white")
```
```{r}
ggplot(df,aes(a))+ geom_density()
```
```{r}
ggplot(df)+ geom_histogram(aes(a,..density..),fill="blue",color="darkgrey")+ 
  geom_density(aes(a,..density..),color="yellow",size=1.4)+
  geom_rug(aes(a))
```
```{r}
ggplot(df,aes(c, color=c)) + geom_point(stat="count",size=4)
```
```{r}
ggplot(df)+ geom_bar(aes(c))+ coord_flip()
```


```{r}
ggplot(df,aes(d,fill=c))+
  geom_bar(position="dodge") +
  scale_fill_grey()

```

```{r}
ggplot(df,aes(d,a)) + geom_boxplot()
```
```{r}
ggplot(df,aes(a,b)) + geom_boxplot(aes(group=cut_width(a,0.5)),outlier.alpha = 0.1) + geom_jitter(aes(color=d),alpha=.5,width = 0.02,size=.7)
```
## facet plot 
```{r}
ggplot(df,aes(c))+ geom_bar() + facet_grid(d~.)
```
## 2D density plot: 
```{r}
ggplot(df,aes(a,b))+geom_density2d()+ geom_point(size=1)
```
```{r}
ggplot(df,aes(a,b)) +
  geom_hex( bins=30)
```

```{r}
ggplot(df,aes(a,b))+ geom_point()+geom_smooth(method=lm)
```
```{r}
a <- ggplot()+geom_point(data=df,aes(a,b,colour=c))
a+theme_bw()
theme_set(theme_bw())

a + theme_grey()
theme_get()
```

```{r}
ggplot(df,aes(a,b))+ 
  geom_point(data=df,aes(group=cut_width(a,0.5)),outlier.alpha=0.1)+
  geom_jitter(width=0.02,size=.7,alpha=.5,color="red")
```


for 3D plots: rayshader

ggplot helper with geoms


```{r}
# combining different plots:
library(patchwork)
?patchwork


p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))

p1 + p2

ggplot(mtcars) +
  geom_point(aes(mpg, disp)) +
  ggplot(mtcars) +
  geom_boxplot(aes(gear, disp, group = gear)) +
  plot_layout(ncol=1,nrow=2,guides="auto") +
  plot_annotation(title="mt cars",subtitle="graphics",tag_levels = 1)

plot_annotation()

```

```{r}
labels <- rownames(mtcars)
ggplot(mtcars) +
  geom_point(aes(mpg,disp,fill=gear)) +
  geom_smooth(aes(mpg,disp),method=lm) + 
  plot_annotation(title="mt cars - data",subtitle="visualization with ggplot",tag_levels = 1)
  
```

## back to Steigerwald Data: 
```{r}
df <- read.csv("https://raw.githubusercontent.com/wegmann/R_data/master/Steigerwald_sample_points_all_data_subset_withNames.csv") 
names(df)
head(df)
```

```{r}
ggplot(df,aes(x=L8.ndvi,y=L8.savi))+
  geom_point() + 
  geom_smooth(method="gam") +
  facet_wrap("LCname") +
  geom_jitter() + 
  geom_violin()


```
```{r}
ggplot(df,aes(x=LCname,y=L8.savi))+
  geom_jitter(aes(alpha=SRTM,size=TimeScan.NDVIsd,color=L8.ndvi)) +
  geom_boxplot(alpha=.5)
```

```{r}
ggplot(df,aes(x=L8.ndvi,y=L8.savi)) + 
    geom_boxplot(aes(fill=LCname),alpha=.3) +
  geom_point(aes(color=LCname),size=2,alpha=.5) +
  geom_smooth()
  facet_grid(.~LCname) 
  
```


-> present a nice plot for next week! 


## get Steigerwald Data into spatial data frame
should be transformed to spatial point data frame(spdf)
of course corresponding x and y coordinates are needed 
```{r}
spdf.obj <- df
names(spdf.obj)

library(sp)

coordinates(spdf.obj) <- ~x+y

library(rgdal)

proj4string(spdf.obj) <- CRS("+init=epsg:32632")



plot(spdf.obj)

spdf.obj@data <- df
class(spdf.obj)

spdf.obj <- spdf.obj



writeOGR(spdf.obj,"sample_point_with_data.shp",driver="ESRI Shapefile","data")

```



